\documentclass[11pt]{article}

% NOTE: The "Edit" sections are changed for each assignment

% Edit these commands for each assignment

\newcommand{\assignmentduedate}{November 16}
\newcommand{\assignmentassignedate}{November 9}
\newcommand{\assignmentnumber}{Ten}

\newcommand{\labyear}{2017}
\newcommand{\labday}{Thursday}
\newcommand{\labtime}{2:30 pm}

\newcommand{\assigneddate}{Assigned: \labday, \assignmentassignedate, \labyear{} at \labtime{}}
\newcommand{\duedate}{Due: \labday, \assignmentduedate, \labyear{} at \labtime{}}

% Edit these commands to give the name to the main program

\newcommand{\mainprogram}{\lstinline{MandelbrotBlackAndWhite}}
\newcommand{\mainprogramsource}{\lstinline{src/main/java/labten/MandelbrotBlackAndWhite.java}}

\newcommand{\secondprogram}{\lstinline{MandelbrotColor}}
\newcommand{\secondprogramsource}{\lstinline{src/main/java/labten/MandelbrotColor.java}}

% Edit this commands to describe key deliverables

\newcommand{\reflection}{\lstinline{writing/reflection.md}}
\newcommand{\readme}{\lstinline{README.md}}

% Commands to describe key development tasks

% --> Running gatorgrader.sh
\newcommand{\gatorgraderstart}{\command{./gatorgrader.sh --start}}
\newcommand{\gatorgradercheck}{\command{./gatorgrader.sh --check}}

% --> Compiling and running program with gradle
\newcommand{\gradlebuild}{\command{gradle build}}
\newcommand{\gradlerun}{\command{gradle run}}

% Commands to describe key git tasks

\newcommand{\gitcommitfile}[1]{\command{git commit #1}}
\newcommand{\gitaddfile}[1]{\command{git add #1}}

\newcommand{\gitadd}{\command{git add}}
\newcommand{\gitcommit}{\command{git commit}}
\newcommand{\gitpush}{\command{git push}}
\newcommand{\gitpull}{\command{git pull}}

\newcommand{\gitaddmainprogram}{\command{git add src/main/java/labten/MandelbrotColorManager.java}}
\newcommand{\gitcommitmainprogram}{\command{git commit src/main/java/labten/MandelbrotColorManager.java -m "Your
descriptive commit message"}}

% Use this when displaying a new command

\newcommand{\command}[1]{``\lstinline{#1}''}
\newcommand{\program}[1]{\lstinline{#1}}
\newcommand{\url}[1]{\lstinline{#1}}
\newcommand{\channel}[1]{\lstinline{#1}}
\newcommand{\option}[1]{``{#1}''}
\newcommand{\step}[1]{``{#1}''}

\usepackage{pifont}
\newcommand{\checkmark}{\ding{51}}
\newcommand{\naughtmark}{\ding{55}}

\usepackage{listings}
\lstset{
  basicstyle=\small\ttfamily,
  columns=flexible,
  breaklines=true
}

\usepackage{fancyhdr}
\usepackage{fancyvrb}

\usepackage[margin=1in]{geometry}
\usepackage{fancyhdr}

\pagestyle{fancy}

\fancyhf{}
\rhead{Computer Science 111}
\lhead{Laboratory Assignment \assignmentnumber{}}
\rfoot{Page \thepage}
\lfoot{\duedate}

\usepackage{titlesec}
\titlespacing\section{0pt}{6pt plus 4pt minus 2pt}{4pt plus 2pt minus 2pt}

\newcommand{\labtitle}[1]
{
  \begin{center}
    \begin{center}
      \bf
      CMPSC 111\\Introduction to Computer Science I\\
      Fall 2017\\
      \medskip
    \end{center}
    \bf
    #1
  \end{center}
}

\begin{document}

\thispagestyle{empty}

\labtitle{Laboratory \assignmentnumber{} \\ \assigneddate{} \\ \duedate{}}

\section*{Objectives}

To further improve your expertise with designing, implementing, and enhancing Java methods, including the completion of
tasks such as creating and calling methods that use boolean expressions, complex conditional logic, and iteration
constructs. Additionally, to practice using Java methods that read input from a text file with the {\tt
java.util.Scanner}. Finally, to fully design and implement a real-world program that provides useful features and to
consider releasing it publicly through GitHub. Since you will complete this assignment with a partner, the last
objective for this project is to further hone your team-work skills and better your ability to use a shared GitHub
repository.

\section*{Suggestions for Success}

\begin{itemize}
  \setlength{\itemsep}{0pt}

\item {\bf Use the laboratory computers}. The computers in this laboratory feature specialized software for completing
  this course's laboratory and practical assignments. If it is necessary for you to work on a different machine, be sure
  to regularly transfer your work to a laboratory machine so that you can check its correctness. If you cannot use a
  laboratory computer and you need help with the configuration of your own laptop, then please carefully explain its
  setup to a teaching assistant or the course instructor when you are asking questions.

\item {\bf Follow each step carefully}. Slowly read each sentence in the assignment sheet, making sure that you
  precisely follow each instruction. Take notes about each step that you attempt, recording your questions and ideas and
  the challenges that you faced. If you are stuck, then please tell a teaching assistant or instructor what assignment
  step you recently completed.

\item {\bf Regularly ask and answer questions}. Please log into Slack at the start of a laboratory or practical session
  and then join the appropriate channel. If you have a question about one of the steps in an assignment, then you can
  post it to the designated channel. Or, you can ask a student sitting next to you or talk with a teaching assistant or
  the course instructor.

\item {\bf Store your files in GitHub}. As in the past laboratory assignments, you will be responsible for storing all
  of your files (e.g., Java source code and Markdown-based writing) in a Git repository using GitHub Classroom. Please
  verify that you have saved your source code in your Git repository by using \command{git status} to ensure that
  everything is updated. You can see if your assignment submission meets the established correctness requirements by
  using the provided checking tools on your local computer and by checking the commits in GitHub.

\item {\bf Keep all of your files}. Don't delete your programs, output files, and written reports after you submit them
  through GitHub; you will need them again when you study for the quizzes and examinations and work on the other
  laboratory, practical, and final project assignments.

\item {\bf Back up your files regularly}. All of your files are regularly backed-up to the servers in the Department of
  Computer Science and, if you commit your files regularly, stored on GitHub. However, you may want to use a flash
  drive, Google Drive, or your favorite backup method to keep an extra copy of your files on reserve. In the event of
  any type of system failure, you are responsible for ensuring that you have access to a recent backup copy of all your
  files.

\item {\bf Explore teamwork and technologies}. While certain aspects of the laboratory assignments will be challenging
  for you, each part is designed to give you the opportunity to learn both fundamental concepts in the field of computer
  science and explore advanced technologies that are commonly employed at a wide variety of companies. To explore and
  develop new ideas, you should regularly communicate with your team and/or the teaching assistants and tutors.

\item {\bf Hone your technical writing skills}. Computer science assignments require to you write technical
  documentation and descriptions of your experiences when completing each task. Take extra care to ensure that your
  writing is interesting and both grammatically and technically correct, remembering that computer scientists must
  effectively communicate and collaborate with their team members and the tutors, teaching assistants, and course
  instructor.

% \item {\bf Review the Honor Code on the syllabus}. While you may discuss your assignments with others, copying source
%   code or writing is a violation of Allegheny College's Honor Code.

\end{itemize}

\section*{Reading Assignment}

To continue to learn more about ``{\tt if/else-if/else}'' statements and boolean expressions, please again review
Sections 5.1--5.3. Since this assignment will also require you to continue to use Java classes and methods, you should
once again review Sections 4.1--4.5. To best prepare for the content in this laboratory assignment, you should also
study Sections 5.4--5.6, paying particularly close attention to the material about {\tt while} loops, {\tt break} and
{\tt continue} statements, text file input, {\tt Iterators}, and the {\tt java.util.ArrayList} class. Students who are
not familiar with text-based todo list management tools are encouraged to review the web site \url{http://todotxt.com/}.

\section*{Accessing the Laboratory Assignment on GitHub}

To access the laboratory assignment, you should go into the \channel{\#announcements} channel in our Slack team and find
the announcement that provides a link for it. Now, make sure that the leader of your team also notes your team number
and first copies this link and pastes it into their web browser. Next, the team leader (i.e., the first person in the
group listing on Slack) will create their team with the name \command{Computer-Science-111-Fall-2017-Lab-9-Group-<group
label>} and then accept the laboratory assignment and see that GitHub Classroom created a new GitHub repository for
your team to access the assignment's starting materials and to store the completed version of your assignment. Note that
the team leader will have to type their group and lab details into a text field. For instance, if the team leader was in
the second group then that person would type ``Group 2 for Lab 9'' into the text field. At this point, each additional
member of the team can accept the assignment through GitHub. Please make sure that each of your team members joins the
team to which the instructor assigned them to work. Unless you provide the instructor with documentation of the
extenuating circumstances that you are facing, not accepting the assignment means that you automatically receive a
failing grade for it. Please see the course instructor with any questions.

Before you move to the next step of this assignment, please make sure that you read all of the content on the web site
for your new GitHub repository, paying close attention to the technical details about the commands that you will type
and the output that your program must produce. Now you are ready to download the starting materials to your laboratory
computer. Click the ``Clone or download'' button and, after ensuring that you have selected ``Clone with SSH'', please
copy this command to your clipboard. To enter the correct directory you should now type \command{cd cs111F2017}. Next,
you can type the command \command{ls} and see that there are some files or directories inside of this directory. By
typing \command{git clone} in your terminal and then pasting in the string that you copied from the GitHub site you will
download all of the code for this assignment. For the previous example, a student would run a \command{git clone}
command in the terminal in this fashion:

\begin{lstlisting}
  git clone git@github.com:Allegheny-Computer-Science-111-F2017/computer-science-111-fall-2017-lab-10-group-2-for-lab-10.git
\end{lstlisting}

After this command finishes, you can use \command{cd} to change into the new directory. If you want to \step{go back}
one directory from your current location, then you can type the command \command{cd ..}. Please continue to use the
\command{cd} and \command{ls} commands to explore the files that you automatically downloaded from GitHub. What files
and directories do you see? What do you think is their purpose? Spend some time exploring, sharing your discoveries with
a neighbor and a \mbox{teaching assistant}.

\noindent
In conclusion, here are some points to remember when creating your program that manages a list:

\begin{enumerate}
  \setlength{\itemsep}{0pt}

\item The provide source code is incomplete---make sure that you add all of the needed features.

\item As in past assignments, your program only needs to have one {\tt main} method in one file.

\item Don't forget that you and your team member(s) should all contribute evenly by typing in text and using commands,
  like \command{git push} and \command{git pull}, to transfer files to and from GitHub.

\end{enumerate}

\section*{Summary of the Required Deliverables}

\noindent Students do not need to submit printed source code or technical writing for any assignment in this course.
Instead, this assignment invites you to submit, using GitHub, the following deliverables.

\begin{enumerate}

  \setlength{\itemsep}{0in}

\item Stored in \reflection{}, a three-paragraph reflection on the commands that you typed in \command{gvim} and the
  terminal window. This Markdown-based document should explain the input, output, and behavior of each command and the
  challenges that you confronted when using it. For every challenge that you encountered, please explain your solution
  for it. This document should also explain how your team collaborated to finish the assignment, with each team member
  writing their own paragraph inside of this Markdown file.

\item A commit log in your GitHub repository that clearly shows that the team members effectively collaborated. That is,
  the commit log should show that commits were evenly made by all team members; the photograph of each member should
  appear in the commit log.

\end{enumerate}

\section*{Evaluation of Your Laboratory Assignment}

Using a report that the instructor shares with you through the commit log in GitHub, you will privately received a grade
on this assignment and feedback on your submitted deliverables. Your grade for the assignment will be a function of the
whether or not it was submitted in a timely fashion and if your program received a green \checkmark{} indicating that it
met all of the requirements. Other factors will also influence your final grade on the assignment. In addition to
studying the efficiency and effectiveness of your Java source code, the instructor will also evaluate the accuracy of
both your technical writing and the comments in your source code. If your submission receives a red \naughtmark{}, the
instructor will reduce your grade for the assignment while still considering the regularity with which you committed to
your GitHub repository and the overall quality of your partially completed work. Please see the instructor if you have
questions about the evaluation of this laboratory assignment.

All team members will receive the same baseline grade for the laboratory assignment. If there are extenuating
circumstances in which one or more of the team members do not effectively collaborate to complete this assignment, then
the course instructor will adjust the grade of specific team members so that it is higher or lower than the baseline
grade, as is fair and necessary. Please see the instructor if you do not understand how he assigns grades for
collaborative assignments. Finally, in adherence to the Honor Code, students should only complete this assignment with
their team members. Deliverables (e.g., Java source code or Markdown-based technical writing) that are nearly identical
to the work of outsiders of your team will be taken as evidence of violating the \mbox{Honor Code}. Please see the
course instructor with any questions about this assignment policy.

\end{document}
