\documentclass[11pt]{article}

% NOTE: The "Edit" sections are changed for each assignment

% Edit these commands for each assignment

\newcommand{\assignmentduedate}{September 28}
\newcommand{\assignmentassignedate}{September 21}
\newcommand{\assignmentnumber}{Four}

\newcommand{\labyear}{2017}
\newcommand{\labday}{Thursday}
\newcommand{\labtime}{2:30 pm}

\newcommand{\assigneddate}{Assigned: \labday, \assignmentassignedate, \labyear{} at \labtime{}}
\newcommand{\duedate}{Due: \labday, \assignmentduedate, \labyear{} at \labtime{}}

% Edit these commands to give the name to the main program

\newcommand{\mainprogram}{\lstinline{DisplayGraphicalScene}}
\newcommand{\mainprogramsource}{\lstinline{src/main/java/labfour/DisplayGraphicalScene.java}}
\newcommand{\mainprograminput}{\lstinline{input/tip_inputs.txt}}

% Edit this commands to describe key deliverables

\newcommand{\reflection}{\lstinline{writing/reflection.md}}

% Commands to describe key development tasks

% --> Running gatorgrader.sh
\newcommand{\gatorgraderstart}{\command{./gatorgrader.sh --start}}
\newcommand{\gatorgradercheck}{\command{./gatorgrader.sh --check}}

% --> Compiling and running program with gradle
\newcommand{\gradlebuild}{\command{gradle build}}
\newcommand{\gradlerun}{\command{gradle run}}

% Commands to describe key git tasks

\newcommand{\gitcommitfile}[1]{\command{git commit #1}}
\newcommand{\gitaddfile}[1]{\command{git add #1}}

\newcommand{\gitadd}{\command{git add}}
\newcommand{\gitcommit}{\command{git commit}}
\newcommand{\gitpush}{\command{git push}}
\newcommand{\gitpull}{\command{git pull}}

\newcommand{\gitaddmainprogram}{\command{git add src/main/java/labfour/DisplayGraphicalScene.java}}
\newcommand{\gitcommitmainprogram}{\command{git commit src/main/java/labfour/DisplayGraphicalScene.java -m "Your
descriptive commit message"}}

% Use this when displaying a new command

\newcommand{\command}[1]{``\lstinline{#1}''}
\newcommand{\program}[1]{\lstinline{#1}}
\newcommand{\url}[1]{\lstinline{#1}}
\newcommand{\channel}[1]{\lstinline{#1}}
\newcommand{\option}[1]{``{#1}''}
\newcommand{\step}[1]{``{#1}''}

\usepackage{pifont}
\newcommand{\checkmark}{\ding{51}}
\newcommand{\naughtmark}{\ding{55}}

\usepackage{listings}
\lstset{
  basicstyle=\small\ttfamily,
  columns=flexible,
  breaklines=true
}

\usepackage{fancyhdr}

\usepackage[margin=1in]{geometry}
\usepackage{fancyhdr}

\pagestyle{fancy}

\fancyhf{}
\rhead{Computer Science 111}
\lhead{Laboratory Assignment \assignmentnumber{}}
\rfoot{Page \thepage}
\lfoot{\duedate}

\usepackage{titlesec}
\titlespacing\section{0pt}{6pt plus 4pt minus 2pt}{4pt plus 2pt minus 2pt}

\newcommand{\labtitle}[1]
{
  \begin{center}
    \begin{center}
      \bf
      CMPSC 111\\Introduction to Computer Science I\\
      Fall 2017\\
      \medskip
    \end{center}
    \bf
    #1
  \end{center}
}

\begin{document}

\thispagestyle{empty}

\labtitle{Laboratory \assignmentnumber{} \\ \assigneddate{} \\ \duedate{}}

\section*{Objectives}

To continue to practice using GitHub to access the files for a laboratory assignment. Additionally, to master the use of
the Ubuntu operating system and software development programs such as a ``terminal window'' and the ``GVim text
editor''. You will also continue to practice using Slack to support communication with the teaching assistants and the
course instructor. Next, you will learn how to implement a Java program that creates graphical objects and displays them
on the user's screen. Finally, you will continue to explore the use of methods, variables, and data types.

\section*{Suggestions for Success}

\begin{itemize}
  \setlength{\itemsep}{0pt}

\item {\bf Use the laboratory computers}. The computers in this laboratory feature specialized software for completing
  this course's laboratory and practical assignments. If it is necessary for you to work on a different machine, be sure
  to regularly transfer your work to a laboratory machine so that you can check its correctness. If you cannot use a
  laboratory computer and you need help with the configuration of your own laptop, then please carefully explain its
  setup to a teaching assistant or the course instructor when you are asking questions.

\item {\bf Follow each step carefully}. Slowly read each sentence in the assignment sheet, making sure that you
  precisely follow each instruction. Take notes about each step that you attempt, recording your questions and ideas and
  the challenges that you faced. If you are stuck, then please tell a teaching assistant or instructor what assignment
  step you recently completed.

\item {\bf Regularly ask and answer questions}. Please log into Slack at the start of a laboratory or practical session
  and then join the appropriate channel. If you have a question about one of the steps in an assignment, then you can
  post it to the designated channel. Or, you can ask a student sitting next to you or talk with a teaching assistant or
  the course instructor.

\item {\bf Store your files in GitHub}. Starting with this laboratory assignment, you will be responsible for storing
  all of your files (e.g., Java source code and Markdown-based writing) in a Git repository using GitHub Classroom.
  Please verify that you have saved your source code in your Git repository by using \command{git status} to ensure that
  everything is updated. You can see if your assignment submission meets the established correctness requirements by
  using the provided checking tools on your local computer and in checking the commits in GitHub.

\item {\bf Keep all of your files}. Don't delete your programs, output files, and written reports after you submit them
  through GitHub; you will need them again when you study for the quizzes and examinations and work on the other
  laboratory, practical, and final project assignments.

\item {\bf Back up your files regularly}. All of your files are regularly backed-up to the servers in the Department of
  Computer Science and, if you commit your files regularly, stored on GitHub. However, you may want to use a flash
  drive, Google Drive, or your favorite backup method to keep an extra copy of your files on reserve. In the event of
  any type of system failure, you are responsible for ensuring that you have access to a recent backup copy of all your
  files.

\item {\bf Explore teamwork and technologies}. While certain aspects of the laboratory assignments will be challenging
  for you, each part is designed to give you the opportunity to learn both fundamental concepts in the field of computer
  science and explore advanced technologies that are commonly employed at a wide variety of companies. To explore and
  develop new ideas, you should regularly communicate with your team and/or the teaching assistants and tutors.

\item {\bf Hone your technical writing skills}. Computer science assignments require to you write technical
  documentation and descriptions of your experiences when completing each task. Take extra care to ensure that your
  writing is interesting and both grammatically and technically correct, remembering that computer scientists must
  effectively communicate and collaborate with their team members and the tutors, teaching assistants, and course
  instructor.

\item {\bf Review the Honor Code on the syllabus}. While you may discuss your assignments with others, copying source
  code or writing is a violation of Allegheny College's Honor Code.

\end{itemize}

\section*{Reading Assignment}

If you have not done so already, please read all of the relevant ``GitHub Guides'', available at
\url{https://guides.github.com/}, that explain how to use many of the features that GitHub provides. In particular,
please make sure that you have read guides such as ``Mastering Markdown'' and ``Documenting Your Projects on GitHub'';
each of them will help you to understand how to use both GitHub and GitHub Classroom. To do well on this assignment, you
should also review Chapter 1 and study Chapter 2 in the course textbook, paying close attention to Sections 2.1 through
2.6, Figure 2.7, and Listing 2.8. See the instructor if you have questions these reading assignments.

\section*{Accessing the Laboratory Assignment on GitHub}

To access the laboratory assignment, you should go into the \channel{\#announcements} channel in our Slack team and find
the announcement that provides a link for it. Copy this link and paste it into your web browser. Now, you should accept
the laboratory assignment and see that GitHub Classroom created a new GitHub repository for you to access the
assignment's starting materials and to store the completed version of your assignment. Specifically, to access your new
GitHub repository for this assignment, please click the green ``Accept'' button and then click the link that is prefaced
with the label ``Your assignment has been created here''. If you accepted the assignment and correctly followed these
steps, you should have created a GitHub repository with a name like
``Allegheny-Computer-Science-111-Fall-2017/computer-science-111-fall-2017-lab-4-gkapfham''. Unless you provide the
instructor with documentation of the extenuating circumstances that you are facing, not accepting the assignment means
that you automatically receive a failing grade for it.

Before you move to the next step of this assignment, please make sure that you read all of the content on the web site
for your new GitHub repository, paying close attention to the technical details about the commands that you will type
and the output that your program must produce. Now you are ready to download the starting materials to your laboratory
computer. Click the ``Clone or download'' button and, after ensuring that you have selected ``Clone with SSH'', please
copy this command to your clipboard. To enter into the right directory you should now type \command{cd cs111F2017}.
Next, you can type the command \command{ls} and see that there are some files or directories inside of this directory.
By typing \command{git clone} in your terminal and then pasting in the string that you copied from the GitHub site you
will download all of the code for this assignment. For instance, if the course instructor ran the \command{git clone}
command in the terminal, it would look like:

\begin{lstlisting}
  git clone https://github.com/Allegheny-Computer-Science-111-F2017/computer-science-111-fall-2017-lab-4-gkapfham.git
\end{lstlisting}

After this command finishes, you can use \command{cd} to change into the new directory. If you want to \step{go back}
one directory from your current location, then you can type the command \command{cd ..}. Please continue to use the
\command{cd} and \command{ls} commands to explore the files that you automatically downloaded from GitHub. What files
and directories do you see? What do you think is their purpose? Spend some time exploring, sharing your discoveries with
a neighbor and a \mbox{teaching assistant}.

\section*{Reading from a File and Computing with Variables}

In order view the source code of a Java program or a Markdown-based writing assignment, you need a text editor. There
are many different text editors on your workstation and you should feel free to explore these on your own. Since it is a
powerful text editor known for helping computer scientists ``edit text at the speed of thought'', in this class we will
normally use the text editor called \command{gvim}. Today, you will view and improve another Java program in {\tt gvim}!
To load the source code of today's Java program into \program{gvim}, you should type \command{gvim
src/main/java/labfour/DisplayGraphicalScene.java} in your terminal window. This should cause a new window, the \program{gvim}
text editor, to appear on your screen. Since \program{gvim} is a modal editor, you may notice that if you start typing,
nothing appears (unless you happen to hit certain letters such as \command{i}, \command{o}, \command{a}, and a few
others). This is because you are not in ``insert mode.'' To get into insert mode, just type the letter \command{i}
(lower case). Once you do this, the window should show \command{--INSERT--} in the lower left corner.

This assignment invites you to implement a Java program that displays exactly ten lines of output. The first
output line should display your name and then today's date. You can accomplish the first task by typing
\command{System.out.println("Gregory M. Kapfhammer " + new Date());}, substituting your name for the name of the course
instructor. Your program will produce nine additional lines of output once you add eight more
\command{System.out.println} statements. Unlike your previous Java programs, this Java program must read input values
from a file and then use these inputs to perform computations and to produce outputs. Please see the Java source code
comments for more details! The remainder of this section summarizes the purpose of the new program.

This assignment asks you to write a Java program, \mainprogramsource{}, that will calculate the tip, the total bill for
the user, and each person's share of the restaurant bill (if there are two or more people). In particular, your program
needs to do the following:

\vspace*{-.1in}
\begin{enumerate}

  \item Ask the user to enter a name (remember to save the user's input into a variable so that it can be used later in
    the program and in the program's output).

  \item Using this name, display a friendly and appropriate welcome message to the user.

  \item Ask the user to enter the restaurant's bill amount (remember to save the user's input into an appropriate
    variable). You should allow for floating point (e.g., decimal or fractional) values.

  \item Ask the user to enter the desired tip percentage as a number between $0$ and $100$ or as a decimal number between
    $0$ and $1$. You have to decide which range you want to use for your program and specify it when prompting for the
    user's input (after picking the correct data type for the percentage, remember to save the user's input into an
    appropriate variable).

  \item Calculate the tip as $tip = \frac{percentage}{100} \times bill$ if the percentage ranges between $0$ and $100$ or
    as $tip = percentage \times bill$ if the percentage ranges between $0$ and $1$.

  \item Calculate the total bill as $total\_bill = bill+tip$

  \item Display to the user:
    \begin{itemize}
      \item The original bill (before the tip)
      \item The tip amount
      \item The total bill (including the tip)
    \end{itemize}

  \item Ask the user how many people will be splitting the total bill (remember to save the user's input into a variable
    so that it can be used later in the program and in the program's output).

  \item Calculate each person's share. For example, if you saved the number of people splitting the bill into a variable
    called {\em num\_people}, then you would calculate each person's share as $share = \frac{total\_bill}{num\_people}$. You should think (or rethink) your views concerning the data types you are using and whether a data
    conversion is required at this point in your program.

  \item Display to the user an exit message that is suitable for an academic setting.

\end{enumerate}

\noindent If a steps says to ``ask the user'' this means the program should read from a file. A sample run is:

\begin{verbatim}
  Gregory M. Kapfhammer Thu Sep 14 12:50:07 EDT 2017
  Timothy, welcome to the tip calculator.
  The amount of your bill is: $50.0
  The percentage you want to tip is: 15.0%
  Alright, that means your total tip will be: $7.5
  Timothy, your total bill is: $57.5
  Okay, I see that there are 2 people in your group.
  So, that means each person must pay: $28.75
  Thanks for using the tip calculator.
  Have an awesome day.
\end{verbatim}

With the exception of the name and date on which the program was run, your \mainprogramsource{} should produce the same
output as seen in the above example. Additionally, you should try to edit the \mainprograminput{} file so that it
provides different inputs to the program. Does your \mainprogram{} still produce the expected output?

Whenever you are finished typing text Java code in \program{gvim}, press the ESC key located in the upper left corner of
the keyboard. This should remove the word \command{--INSERT--} from the bottom of the screen and take you out of insert
mode. Use the \option{File/Save} menu to save your program. Alternatively, if you would like to use the keyboard to save
your file, you can press \command{:w} when you are not in insert mode. Leaving the {\tt gvim} window open, go back to
your terminal window.

If you want to \step{build} your program you can type the command \gradlebuild{} in your terminal, thereby causing the
Java compiler to check your program for errors and get it ready to run. If you get any error messages, go back into
\program{gvim} and try to figure out what you mis-typed and fix it. Once you have solved the problem, make a note of the
error and the solution for resolving it. Re-save your program and then build it again by re-running the \gradlebuild{}.
If you cannot build \mainprogram{} correctly, then please talk with a teaching assistant or the instructor.

When all of the errors are eliminated, you can run your program by typing \gradlerun{} in the terminal window---this is
the ``execute'' step that will run your program and produce the designated output. You should see your name, today's
date, and nine more lines of text. Make sure there are spaces separating words in your output (did you forget to put a
space inside the quotation marks after your last name?). If not, then repair the program and re-build and re-run it.
Once the program runs, please reflect on this process. What step did you find to be the most challenging? Why? You
should write your reflections in a file, called \reflection{}, that uses the Markdown writing language. To complete this
aspect of the assignment, you should write one high-quality paragraph that reports on your experiences with the various
commands and Java code segments.

\section*{Checking the Correctness of Your Program and Writing}

As verified by the Checkstyle tool, the source code for the \mainprogram{} file must adhere to all of the requirements
in the Google Java Style Guide available at \url{https://google.github.io/styleguide/javaguide.html}. The Markdown file
that contains your reflection must adhere to the standards described in the Markdown Syntax Guide
\url{https://guides.github.com/features/mastering-markdown/}. Finally, your \reflection{} file should adhere to the
Markdown standards established by the \step{Markdown linting} tool available at
\url{https://github.com/markdownlint/markdownlint/} and the writing standards set by the \step{prose linting} tool from
\url{http://proselint.com/}. Instead of requiring you to manually check that your deliverables adhere to these
industry-accepted standards, the GatorGrader tool that you will use in this laboratory assignment makes it easy for you
to automatically check if your submission meets the standards for correctness.

Since this is not your first laboratory assignment, you will notice that the provide source code does not contain all of
the required comments at the top of the Java source code file. This means that you will have to inspect the source code
from previous laboratory and practical assignments to review how to create the comments in the \mainprogramsource{}
file. Moreover, the provided source code is missing many of the lines that are needed to pass the GatorGrader checks.
For instance, GatorGrader will check to ensure that \mainprogram{} produces exactly ten lines of output, that you use
the \command{new Date()} construct in the Java source code, and that the program correctly produces the same output as
seen on the previous page.

To get started with the use of GatorGrader, type the command \gatorgraderstart{} in your terminal window. Once this step
completes you can type \gatorgradercheck{}. If your work does not meet all of the assignment's requirements, then you
will see the following output in your terminal: \command{Overall, are there any mistakes in the assignment? Yes}. If you
do have mistakes in your assignment, then you will need to review GatorGrader's output, find the mistake, and try to fix
it. Once your program is building correctly, fulfilling at least some of the assignment's requirements, you should
transfer your files to GitHub using the \gitcommit{} and \gitpush{} commands. For example, if you want to signal that
the \mainprogramsource{} file has been changed and is ready for transfer to GitHub you would first type
\gitcommitmainprogram{} in your terminal, followed by typing \gitpush{}, and then checking to see that the transfer to
GitHub is successful. If you notice that the network communication with GitHub did not work correctly, then please try
to determine why, asking a teaching assistant or the instructor for assistance, if necessary.

After the course instructor enables \step{continuous integration} with a system called Travis CI, when you use the
\gitpush{} command to transfer your source code to your GitHub repository, Travis CI will initialize a \step{build} of
your assignment, checking to see if it meets all of the requirements. If both your source code and writing meet all of
the established requirements, then you will see a green \checkmark{} in the listing of commits in GitHub after awhile.
If your submission does not meet the requirements, a red \naughtmark{} will appear instead. The instructor will reduce a
student's grade for this assignment if the red \naughtmark{} appears on the last commit in GitHub immediately before the
assignment's due date. Yet, if the green \checkmark{} appears on the last commit in your GitHub repository, then you
satisfied all of the main checks, thereby allowing the course instructor to evaluate other aspects of your source code
and writing, as further described in the \step{Evaluation} section of this assignment sheet. Unless you provide the
instructor with documentation of the extenuating circumstances that you are facing, no late work will be considered
towards your grade for this laboratory assignment.

In conclusion, here are some points to remember as you complete this laboratory assignment:

\begin{enumerate}
  \setlength{\itemsep}{0pt}

\item The provided source code uses the {\tt Scanner} class; notice that it is imported at the top of the program. You
  may refer to the textbook for code demonstrating the use of the {\tt Scanner}; you need to use methods that support
  the input for variables of type {\tt String}, {\tt int}, and {\tt double}. Picking the wrong data types for your
  variables may cause the program to crash or to produce incorrect output. See the instructor or an assistant if you are
  not sure what types to use.

\item Students who want to learn more about how to use the {\tt Scanner} class can refer to pages 90 and 91 of the
  textbook, focusing on the {\tt GasMileage} class provided in Listing 2.9.

\item You should think carefully about what data type you want to use for your variables.

\item As in past assignments, your program only needs to have one {\tt main} method.

\item To display a dollar sign, you can just type ``\$'' inside the string of your output statement.

\item Note that your program will alternate between printing and computing---this
is okay!

\item Don't forget to review the assignment sheets from the previous laboratory and practical assignments as they
  contain insights that will support your completion of this assignment.

\end{enumerate}


\section*{Summary of the Required Deliverables}

\noindent Students do not need to submit printed source code or technical writing for any assignment in this course.
Instead, this assignment invites you to submit, using GitHub, the following deliverables.

\begin{enumerate}

  \setlength{\itemsep}{0in}

\item Stored in \reflection{}, a one-paragraph reflection on the commands that you typed in \command{gvim} and the
  terminal window. This Markdown-based document should explain the input, output, and behavior of each command and the
  challenges that you confronted when using it. For every challenge that you encountered, please explain your solution
  for it.

\item A properly documented, well-formatted, and correct version of \mainprogramsource{} that both meets all of the
  established requirements and produces the desired output. In particular, please recall that this Java program should
  read inputs from a file, perform computations, and display the correct outputs in the terminal window.

\end{enumerate}

\section*{Evaluation of Your Laboratory Assignment}

Using a report that the instructor shares with you through the commit log in GitHub, you will privately received a grade
on this assignment and feedback on your submitted deliverables. Your grade for the assignment will be a function of the
whether or not it was submitted in a timely fashion and if your program received a green \checkmark{} indicating that it
met all of the requirements. Other factors will also influence your final grade on the assignment. In addition to
studying the efficiency and effectiveness of your Java source code, the instructor will also evaluate the accuracy of
both your technical writing and the comments in your source code. If your submission receives a red \naughtmark{}, the
instructor will reduce your grade for the assignment while still considering the regularity with which you committed to
your GitHub repository and the overall quality of your partially completed work. Please see the instructor if you have
questions about the evaluation of this laboratory assignment.

% Please note that at the time this assignment was released, the Travis CI servers were not correctly configured for
% testing and thus you will not immediately be able to observe the \checkmark{} or \naughtmark{} in your commit listing on
% GitHub. Once this service again becomes available, you will be able to complete this step of the assignment, if you
% decide to do so. However, you can receive full credit for the assignment as long as your source code and writing pass
% the local checks run by typing \gatorgradercheck{}. Please see the instructor if you have questions about this issue.

\section*{Adhering to the Honor Code}

In adherence to the Honor Code, students should complete this assignment on an individual basis. While it is appropriate
for students in this class to have high-level conversations about the assignment, it is necessary to distinguish
carefully between the student who discusses the principles underlying a problem with others and the student who produces
assignments that are identical to, or merely variations on, someone else's work. Deliverables (e.g., Java source code or
Markdown-based technical writing) that are nearly identical to the work of others will be taken as evidence of violating
the \mbox{Honor Code}. Please see the course instructor if you have questions about this policy.

\end{document}
